---
title: "The Top 100 Young Adult Books"
subtitle: "Some Examples"
author: "Robyn Steveley"
date: "2024-02-18"
categories: [other, notes,]

code-fold: true
code-tools: true
description: "Examining the Data in the File"
draft: true
---

## Set Up

```{r}
#| output: false
library(tidyverse)
#library(ggplot2)
```

```{r}
BOOKS <- read.csv("goodreads_Top100_YoungAdultFiction1.csv")
```

## The Graphs

An overview of the length of these books. A Histogram.

```{r}
#Book length (number of pages)
fig <- ggplot(BOOKS, aes(pages))+ geom_histogram(binwidth = 20, fill="blue")

#Style
fig_labs <- labs(title = "Number of Pages for Top 100 Young Adult Books")

fig_theme <- theme(panel.background=element_rect(fill="beige"))

#Showing figure 4
fig <- fig + fig_labs + fig_theme

print(fig)
```

Series vs Non-Series. A bar plot... maybe?

I would like to figure out how to make this with two bars, one for series and one for stand-alones. Or I would like to change the labels on the bars.

```{r}
#Series vs non-series
#BOOKS$series_Q <- ifelse(is.na(BOOKS$series), "Stand Alone", "Series") 

fig <- ggplot(BOOKS, aes(x = numberOfSeries))+ geom_bar(position = "dodge", color = "navy", fill = "blue")

#Style
fig_label <- labs(title = "Books in Series Vs. Stand Alones")
fig_theme <- theme(panel.background=element_rect(fill="white"))

#Showing figure 4
fig <- fig + fig_label + fig_theme

print(fig)
```

Number of Books per Genre. An exploration of complex data tables. (In progress...)

```{r}

```

How Popular of these books?

A histogram of the number of books with a certain rating and a scatter plot of ratings vs. number of ratings.

```{r}
#Number of Ratings
fig <- ggplot(BOOKS, aes(x = numRatings))+ geom_histogram(binwidth = 50000)

#Style
fig_labs <- labs(title = "Number of Ratings for Top 100 Young Adult Books")

fig_theme <- theme(panel.background=element_rect(fill="white"))

#Showing figure 4
fig <- fig + fig_labs + fig_theme

print(fig)
```

```{r}
#Number of Ratings vs. Rating
fig <- ggplot(BOOKS, aes(x = numRatings, y = rating))+ geom_point()

#Style
fig_labs <- labs(title = "Number of Ratings for Top 100 Young Adult Books")

fig_theme <- theme(panel.background=element_rect(fill="white"))

#Showing figure 4
fig <- fig + fig_labs + fig_theme

print(fig)
```

Multiple Authors? Yes? No? (do I even want to try to make this?)

```{r}

```

Setting? An exploration of searching.

```{r}

```

Book Format

```{r}
fig <- ggplot(BOOKS, aes(x = bookFormat))+ geom_bar(position = "dodge", color = "red", fill = "pink")

#Style
fig_label <- labs(title = "Books format: a completly useless graph")
fig_theme <- theme(panel.background=element_rect(fill="white"))

#Showing figure 4
fig <- fig + fig_label + fig_theme

print(fig)
```

Book Format: series vs non series

Pages by genre

Publisher?

```{r}

```

Genre by Publisher

Year published

```{r}
#Number of Ratings
fig <- ggplot(BOOKS, aes(x = firstPublishYear))+ geom_histogram(binwidth = 1)

#Style
fig_labs <- labs(title = "Number of Ratings for Top 100 Young Adult Books")

fig_theme <- theme(panel.background=element_rect(fill="white"))

#Showing figure 4
fig <- fig + fig_labs + fig_theme

print(fig)
```

Rating Per Year

```{r}
#Number of Ratings vs. Rating
fig <- ggplot(BOOKS, aes(x = firstPublishYear, y = numRatings))+ geom_point()

#Style
fig_labs <- labs(title = "Number of Ratings for Top 100 Young Adult Books")

fig_theme <- theme(panel.background=element_rect(fill="white"))

#Showing figure 4
fig <- fig + fig_labs + fig_theme

print(fig)
```

First year published vs year published

```{r}

```
